---
title: "PAR_Regional_Variation"
author: "Becca Cates"
date: "2026-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# currently there are two datasets, one is a "historic" or "science quality" dataset that goes through September 2022, and the other is a NRT "real time" dataset which has more recent data. NRT is not up to date as they QA/QC data before making publicly available. Looks it is available through beginning of 2024.

```{r}
install.packages("sf")
install.packages("ggspatial")
install.packages("geosphere")
install.packages("data.table")
install.packages("prettymapr")
install.packages("raster")

require(sf)
require(tidyverse)
require(ggspatial)
require(geosphere)
require(data.table)
```

#Farm info
```{r}
# site coords from proposal

SBOlat <- 60.65912
SBOlon <- -145.88357
ROKlat <- 60.56503
ROKlon <- -145.9696
SBRlat <- 60.6343
SBRlon <- -145.99634
SSFlat <- 59.460476
SSFlon <- -151.51823
MIOlat <- 59.5702
MIOlon <- -151.2695
BCFlat <- 59.46815
BCFlon <- -151.5175
KOBlat <- 57.53253
KOBlon <- -154.0354
KISlat <- 57.76642
KISlon <- -152.4116
AOFlat <- 57.65823
AOFlon <- -152.4205


SBO <- st_as_sf(data.frame(lat = SBOlat, lon = SBOlon), coords = c("lon", "lat"), crs = "epsg:4326")
ROK <- st_as_sf(data.frame(lat = ROKlat, lon = ROKlon), coords = c("lon", "lat"), crs = "epsg:4326")
SBR <- st_as_sf(data.frame(lat = SBRlat, lon = SBRlon), coords = c("lon", "lat"), crs = "epsg:4326")
SSF <- st_as_sf(data.frame(lat = SSFlat, lon = SSFlon), coords = c("lon", "lat"), crs = "epsg:4326")
MIO <- st_as_sf(data.frame(lat = MIOlat, lon = MIOlon), coords = c("lon", "lat"), crs = "epsg:4326")
BCF <- st_as_sf(data.frame(lat = BCFlat, lon = BCFlon), coords = c("lon", "lat"), crs = "epsg:4326")
KOB <- st_as_sf(data.frame(lat = KOBlat, lon = KOBlon), coords = c("lon", "lat"), crs = "epsg:4326")
KIS <- st_as_sf(data.frame(lat = KISlat, lon = KISlon), coords = c("lon", "lat"), crs = "epsg:4326")
AOF <- st_as_sf(data.frame(lat = AOFlat, lon = AOFlon), coords = c("lon", "lat"), crs = "epsg:4326")

# map service 

esri_ocean <- paste0('https://services.arcgisonline.com/arcgis/rest/services/',
                     'Ocean/World_Ocean_Base/MapServer/tile/${z}/${y}/${x}.jpeg')

```

#Bounding Box - SBO
```{r}
# helper function for creating bounding box
getsquare <- function(x) {
  # d = distance in meters, here using 15km
    as.data.frame(destPoint(c(SBOlon, SBOlat), b = x, d = 20000)) #need to recode for farm here
}

# bound box 
bbox <- st_bbox(st_as_sf(map_df(c(45,135,225,315), getsquare), coords = c("lon", "lat"), crs = "epsg:4326"))

bboxRect <- st_as_sfc(bbox)

mapfig <- ggplot(data = bboxRect) +
  geom_sf(fill = NA) +
  annotation_map_tile(type = esri_ocean, zoom = 12) +
  geom_sf(data = bboxRect, fill = "grey20", alpha = 0.2) +
  geom_sf(data = SBO) ####need to recode for farm here

mapfig
#ggsave("figures/map.png", mapfig)

#### SQ data (2002-2023) ####

base_url <- "https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1par08day.csvp?par%5B(2003-01-05T00:00:00Z):1:(2022-06-14T00:00:00Z)%5D%5B(MAXLAT):1:(MINLAT)%5D%5B(MINLON):1:(MAXLON)%5D"

url <- str_replace(base_url,pattern = "MINLAT", as.character(round(bbox$ymin, 5))) |>
  str_replace(pattern = "MAXLAT", as.character(round(bbox$ymax, 5))) |>
  str_replace(pattern = "MAXLON", as.character(round(bbox$xmax, 5))) |>
  str_replace(pattern = "MINLON", as.character(round(bbox$xmin, 5))) 

pardata <- data.table::fread(url)

colnames(pardata) <- c("datetime", "latitude", "longitude", "par")

historic <- pardata |>
  mutate(date = as.Date(datetime)) |>
  group_by(date) |>
  summarise(par = mean(par, na.rm=TRUE)) |>
  mutate(source = "sq")


#### RT data (2002-2024) ####

base_url_rt <- "https://coastwatch.pfeg.noaa.gov/erddap/griddap/erdMH1par08day_R2022NRT.csvp?par%5B(2022-03-02):1:(2024-01-29T00:00:00Z)%5D%5B(MAXLAT):1:(MINLAT)%5D%5B(MINLON):1:(MAXLON)%5D"

url_rt <- str_replace(base_url_rt,pattern = "MINLAT", as.character(round(bbox$ymin, 5))) |>
  str_replace(pattern = "MAXLAT", as.character(round(bbox$ymax, 5))) |>
  str_replace(pattern = "MAXLON", as.character(round(bbox$xmax, 5))) |>
  str_replace(pattern = "MINLON", as.character(round(bbox$xmin, 5))) 

pardatart <- data.table::fread(url_rt)

write.csv(pardata, "Data/pardata.csv")

colnames(pardatart) <- c("datetime", "latitude", "longitude", "par")

realtime <- pardatart |>
  mutate(date = as.Date(datetime)) |>
  group_by(date) |>
  summarise(par = mean(par, na.rm=TRUE)) |>
  mutate(source = "rt")

# combined

alldata <- rbind(historic, realtime)

parfig <- alldata |>
  filter(date > as.Date("2020-01-01")) |> ###can adjust date here
  ggplot(aes(x = date, y = par, color = source)) +
  geom_line() +
  theme_classic() +
  labs(x = "", y = "PAR (einstein m-2 day-1)")

parfig

#ggsave("figures/par.png", parfig)
```

##### Copy and paste above code and recode for each farm site or create loop to apply to all farms

#Additional code below used for SPM if you would like that for EVOS farm sites - relies on same bbox lat long defined above for PAR 
#### Suspended particulate matter ####
#Note this is a proxy for SPM - its actually remote sensing reflectance (Rrs (λ)) per steradian (sr-1) at 671 nm. 
```{r}
base_spm_url <- "https://coastwatch.pfeg.noaa.gov/erddap/griddap/nesdisVHNSQnLw671Weekly.csvp?nLw_671%5B(YEAR1-01-01T12:00:00Z):1:(YEAR2-01-01T12:00:00Z)%5D%5B(0.0):1:(0.0)%5D%5B(MAXLAT):1:(MINLAT)%5D%5B(MINLON):1:(MAXLON)%5D"

spm_url <- str_replace(base_spm_url,pattern = "MINLAT", as.character(round(bbox$ymin, 5))) |>
  str_replace(pattern = "MAXLAT", as.character(round(bbox$ymax, 5))) |>
  str_replace(pattern = "MAXLON", as.character(round(bbox$xmax, 5))) |>
  str_replace(pattern = "MINLON", as.character(round(bbox$xmin, 5))) 


get_spm <- function() {
  spmyear <- list()
  for(i in 2023:2026) {
    
    spmyear[[i]] <- str_replace(spm_url, pattern = "YEAR1", as.character(i)) |>
      str_replace(pattern = "YEAR2", as.character(i+1)) |>
      fread()
      
  }
  bind_rows(spmyear)
}

spmdata <- get_spm()

write.csv(spmdata, "data/spmdata.csv")

spmavg <- spmdata |>
  mutate(date = as.Date(`time (UTC)`)) |>
  group_by(date) |>
  summarise(spm = mean(`nLw_671 (W m^-2 um^-1 sr^-1)`, na.rm=TRUE)) 

spmfig <- ggplot(spmavg, aes(x = date, y = spm)) +
  geom_line() +
  theme_classic() +
  labs(x = "", y = "nLw_671 (W m^-2 um^-1 sr^-1)")

spmfig

#ggsave("figures/spmfig.png", spmfig)

spmmap <- spmdata |>
  mutate(lat = `latitude (degrees_north)`, lon = `longitude (degrees_east)`, spm = `nLw_671 (W m^-2 um^-1 sr^-1)`) |>
  mutate(date = as.Date(`time (UTC)`)) |>
  group_by(lat, lon) |>
  summarise(spm = mean(spm, na.rm=TRUE)) |>
  st_as_sf(coords = c("lon","lat"), crs = "epsg:4326") |>
  ggplot() +
  annotation_map_tile(type = esri_ocean, zoom = 11) +
  geom_sf(aes(color = spm), size = 12, alpha = 0.6, shape = 15) +
  scale_color_viridis_c() 

spmmap

#ggsave("figures/spmmap.png", spmmap, height = 3, width = 9, dpi=300)

```


