---
title: "EVOS_Regional_Variation_T1_T2"
author: "Becca Cates"
date: "2025-05-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# library
```{r}
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

#read in data
```{r}
oyster <- read.csv("Data/EVOS Oyster morphometrics.csv")
```

#clean and prep data
```{r}
str(oyster)

#create subset of data
oyster <- dplyr::select(oyster, c("Sample_event", "Sample_date", "Region", "Farm", "Oyster_age", "Shell_height_mm", "Shell_length_mm", "Shell_depth_mm", "Whole_weight_g", "Meat_weight_g"))

#change str of data
oyster$Sample_date <- as.Date(oyster$Sample_date, "%m/%d/%Y")
oyster$Meat_weight_g <- as.numeric(oyster$Meat_weight_g)
oyster$Whole_weight_g <- as.numeric(oyster$Whole_weight_g)
oyster$Shell_depth_mm <- as.numeric(oyster$Shell_depth_mm)
oyster$Shell_length_mm <- as.numeric(oyster$Shell_length_mm)
oyster$Sample_event <- as.factor(oyster$Sample_event)

#Create cup ratio column
oyster$cup_ratio <- 1
oyster$cup_ratio <- as.numeric(oyster$cup_ratio)
oyster$cup_ratio <- as.numeric(oyster$Shell_depth_mm)/as.numeric(oyster$Shell_height_mm)

#create subadult adn juvenile dfs
subadult <- filter(oyster, Oyster_age == "Subadult")
juvenile <- filter(oyster, Oyster_age == "Juvenile")
```

#######################Subadults

#Boxplot: Subadult meat weight (g) by farm and sample event  
```{r}
ggplot(subadult, aes(x=Farm, y=Meat_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()

sub_meat <- aov(formula = Meat_weight_g ~ Region + Sample_event, data = subadult)
summary(sub_meat)
TukeyHSD(sub_meat)
```

#Boxplot: Subadult meat weight (g) by region and sample event
```{r}
ggplot(subadult, aes(x=Region, y=Meat_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#Boxplot: Subadult whole weight (g) by region and sample event
```{r}
ggplot(subadult, aes(x=Region, y=Whole_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#Boxplot: Subadult cup ratio by farm and sample event
```{r}
ggplot(subadult, aes(x=Region, y=cup_ratio, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#faceted boxplots
```{r}
# One box per treatment
p1 <- ggplot(subadult, aes(x=Farm, y=Meat_weight_g, fill= Sample_event)) + 
    geom_boxplot() +
  theme_classic2() +
    facet_wrap(~Sample_event)+
  theme(strip.text = element_blank())
p1
# one box per variety
p2 <- ggplot(subadult, aes(x=Farm, y=Meat_weight_g, fill=Sample_event)) + 
    geom_boxplot() +
    facet_wrap(~Farm, scale="free")+
    theme_classic2() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p2
```

#######################Juveniles

#Boxplot: Juvenile meat weight (g) by farm and sample event  
```{r}
ggplot(juvenile, aes(x=Farm, y=Meat_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#Boxplot: Juvenile meat weight (g) by region and sample event
```{r}
ggplot(juvenile, aes(x=Region, y=Meat_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#Boxplot: Juvenile whole weight (g) by region and sample event
```{r}
ggplot(juvenile, aes(x=Region, y=Whole_weight_g, fill= Sample_event)) + 
  geom_boxplot() +
  theme_classic2()
```

#faceted boxplots
```{r}
# One box per treatment
p1 <- ggplot(juvenile, aes(x=Farm, y=Meat_weight_g, fill= Sample_event)) + 
    geom_boxplot() +
  theme_classic2() +
    facet_wrap(~Sample_event)+
  theme(strip.text = element_blank())
p1
# one box per variety
p2 <- ggplot(juvenile, aes(x=Farm, y=Meat_weight_g, fill=Sample_event)) + 
    geom_boxplot() +
    facet_wrap(~Farm, scale="free")+
    theme_classic2() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
p2
```

